\section{Cubedate Implementation}
\label{sec:implementation}

We implemented Cubedate by extending the open source ecosystem around RIOT, which we combined with our initial firmware running on the Thingsat payload, also open sourced, which we published and maintain at XXX. \todoEB{Link to repo here}


\underline{Bus implementation :}
To provide a simple socket-like API on top of the raw CAN bus connecting hosted payloads and OBC on the cubesat, we used \href{https://github.com/libcsp/libcsp}{libCSP}, a rough library implementing the Cubesat Space Protocol (CSP version 1). Roughly, CSP provides an "equivalent" of the IP/UDP stack and static routing. This approach mimics a number of other cubesat designs. In practice, we integrated libCSP in RIOT, to simplify the programming of communication between the OBC and the payload.%More generally, libCSP aims to support the most common serial communication interfaces found on \textit{Satellite} payloads, OBC and components (links such as I2C, RS-232, CAN and wrappers for TAP, ZMQ etc.) 
\todoEB{Here, can we cite other cubesat which run libCSP?}.


\underline{SUIT implementation :}
To fully support the wide variety of software updates case required by Cubedate, we extended existing RIOT implementation of SUIT~\cite{zandberg2019secure} which, initially, supported only firmware updates, internal storage, and WPAN network delivery. We generalized the SUIT state machine to add support for :
% with the extra mechanisms for updating more general software updates, including mission scenario files and more generally arbitrary configuration files. The state machine

\begin{itemize}
    \item \textbf{Heterogeneous update delivery mechanisms}: configurable \{message model, network stack, network interface\} bundle, or FileSystem \textit{read/write} functions.
%    \item \textit{manifest} and data resource to the SUIT state machine.
    \item \textbf{Heterogeneous storage destination}: configurable internal/external memory: Volatile 
     (e.g. RAM for mission files, or for tiny runtime execution containers such as FemtoContainers~\cite{zandberg2021femto})
    or non-volatile (e.g. FileSystem or internal Flash).
     \item \textbf{Heterogeneous update data URI}: either a local file (e.g.: mounted USB device) or a remotely accessible file (eg CoAP or HTTP endpoint);
\end{itemize}

\underline{Pull implementations :}
We implemented several alternatives for the \texttt{client/server} interaction which must be initiated by the hosted payload to pull (GET) software update data over the network from either the OBC, or a (LPWAN) groundstation.
%\francisco{Need some more info on UP, a link?}.

The first pull mechanism we implemented uses CoAP~\cite{rfc7252}, an open standard low-power IPv6 protocol following a general-purpose REST architecture.
%The advantage of using CoAP is that it is general purpose for resource-constrained networked nodes. 
In particular, it is not only applicable over the CAN/CSP link, but also usable over VHF links~\cite{palma2018vhf-coap} or LoRa links~\cite{sanchez2018lora-coap}, which paves the way for some Cubedate software updates over the direct link from a Thingsat LoRa groundstation to the Thingsat payload.

However, Cubedate is agnostic w.r.t the pull mechanism. To demonstrate this, we also implemented an alternative pull mechanism using UP (the Universal Platform protocol), a dedicated proprietary protocol provided by the cubesat operator for \texttt{Request/Response} interaction with the OBC. This option is currently deployed and running on ThingSat in orbit. 

\iffalse

\subsection{Message Model, Request/Response Semantics}

A \texttt{client/sever} model adequately represents the interactions between the SUIT state
machine (ran in the \textit{payload}) and the Data Resource. In the scope of the
update process the \textit{payload} will act as the client requesting a \textit{manifest}
and sequently the \textit{Data Resource}, but in the global picture it will act
as both \texttt{client \& server}, answering to queries from the OBC or the
ground-segment. ThingSat uses UP \francisco{some info on UP}, a proprietary message
model enabling request-response type interactions, which can be replaced by a
generic \texttt{Request/Response} layer.

\texttt{CoAP} is therefore used because its designed to run on a connection-less transport
and offers little overhead in the exchanged messages as well as support for exchanging
messages larger than the underlying protocols MTU (\francisco{cite block coap}
(4 \texttt{bytes} \texttt{CoAP} header, 0-3 extra \texttt{bytes} for block transmissions).
Using a standard like \texttt{CoAP} allows for simplified integration into \textit{North-bound}
application code, e.g.: \texttt{SUIT} as well as allowing to easily replace the
\texttt{South-bound} network-stack.

Note that a \texttt{Request/Response} approach is not a requirement, but makes
for a clean architecture.

\subsection{Network Stack}

A CubeSat Network will commonly be split into two segments: an in-orbit segment
(OBC \& \textit{payloads}) and a Ground-segment (Ground Station). This motivated
the existence of the \textit{Cubesat Space Protocol CSP} deployed on Satellites
such as Nuts, GomSpace, etc. \francisco{cite}. But recent deployments see CubeSats
as part of larger Network such as LoRaWan \francisco{cite fossa}, and IP based
networks are also present on CubeSats.

\textit{CSP} usage was a requirements specified from SatRevolution as the network
stack to use for ThingSat. Nonetheless this implementation aspired to be agnostic
in design to the network stack, allowing to easily replace it by e.g. an
\texttt{IPV6/UDP} or \texttt{LoRaWAN} stack.

This was done by using a {\bf A generic North-bound interface} called \texttt{sock}
which provides a socket-like API for network stacks. Although designed for IP,
it can be wrapped around CSP if care is taken to handle addressing.

\subsection{Communication Bus}

\texttt{LibCSP} and \textit{CSP} support several physical layers such as CAN, I2C,
RS-232, network interfaces wrappers also exist for TAP, ZMQ, and can be easily
extended to cover the radio link between the \textit{Satellite} and
\textit{Ground-Segment}. The physical layers supported are not random, they match
the most common serial communication interfaces found on \textit{Satellite}
payloads, OBC and components. Ethernet is also commonly available  but prohibitive
in most microcontrollers based use-case because of its high power consumption.
\francisco{add some details on CAN?}

CAN was used since it provides an extremely robust, simple and efficient,
widely proven in the automotive, flight and space industry.

\subsection{Operating System: RIOT}

RIOT is a general-purpose OS designed for small IoT devices based on micro-controllers.
It provides with blah blah..

\subsubsection{Integrating LibCSP and RIOT}

Details on the integrations work, learned lessons and sufferings.


\subsection{Continuous Integration \- Collaborative Platform}

The CubeDate architecture was implemented on ThingSat (using UP instead of CoAP).
A Continuous Integration collaborative platform mimicking the update workflow described
in section \francisco{section?}. This consist on a minimal implementation of the
\texttt{OBC} - \texttt{payload} interactions.

The \texttt{OBC} can be ran on RIOTs \texttt{native} platform with a mounted
FileSystem or on any RIOT \texttt{board} with an `sdcard` slot or adapter.
\textit{manifests} and matching \textit{Data-Resources} can be pushed to the
\texttt{OBC} which in its turn acts as a \texttt{CoAP} file-server.

The Payload can run on RIOTs \texttt{native} platform or on any \texttt{RIOT}
\texttt{board}. Different \textit{Data-Resources} can be delivered to RAM,
External NVM or internal NVM (FLASH), simulating delivery of mission files,
firmware updates or other.

Both \texttt{OBC} and \texttt{payload} run over \texttt{LibCSP} over CAN. If
using \texttt{native} \texttt{SocketCAN} is used to create virtual can
devices.

\francisco{Place Holder of Diagram of the CI}

The developments related to this paper where in release in open-source form for
the benefit of CubeSat and RIOT communities and can be found at.
\fi
