\section{ThingSat}
\label{sec:case-study}

%\subsection{Overview}
\paragraph*{Mission Goal:} in-orbit observation of glaciers in Europe and in Polynesia for academic experimental research purposes.
\paragraph*{Approach}: buy/rent rack space (payload slot) on a low-earth orbit CubeSat. We used STORK, a 3U CubeSat designed and operated by the Polish company SatRevolution, launched January 2022. Such CubeSats offer payload slots ranging from 0,25U slots, to 0,5U slots or a 1U slot. Expected lifetime is 5 years. Launched by SpaceX or Virgin Orbit.
%\paragraph*{High-level overview of Segments}

\subsection{Distributed System Architecture Overview}
\paragraph*{Ground Segment Description}
\paragraph*{Space Segment Description} 0,25U slot (??)
\paragraph*{Control Segment} % Francisco: not sure if its the right name
\paragraph*{OBC \& Payload Communication Bus}
\paragraph*{Payload Description}: tenant status, energy budget, on time etc.
% - OBC -> SatRev controlled
% - Payload -> Full Control

\subsection{Communication Characteristics Overview}
\paragraph*{link-budget Mission Control}: delivering and communicating mission
files, delivering software updates, latency, etc.
\paragraph*{link-budget Mission}: communication between OBC and Payload, payload
active time for communication with ground segments.

\subsection{Software Updates Requirements}
\paragraph*{What needs updates}: ?
\paragraph*{Why?}: timeline for deployment, CoVID context, etc. making it impossible
to deliver a fully functional device in time. Need to update the firmware remotely
and in orbit.
\paragraph*{Communication Chain}: how are new firmware, payloads delivered,
who has access, etc..

